### 2015-12-09 R/testRandomForest.Rmd
Fixex sparse random forrest

Run rf on a small saple of genom
========================================================

Run rf on a small saple of genom

```{r}
library(randomForest)
data <-read.table('wide-data.csv',header=FALSE, sep=",")

X <- t(data[-1,])
y <- as.factor(t(data[1,]))

model <-randomForest(X,y,ntree=100, importance=TRUE)
```

Plot OOB error rates

```{r fig.width=7, fig.height=6}
plot(model)
```

```{r}
model$err.rate[90:100,]
```

Plot variable importance

```{r fig.width=10, fig.height=7}
varImpPlot(model)
```